Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-08 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
More editing of references in Madrzykowski thesis.

Heskestad developed an empirical method for calculating the mean visible flame height for fuels which do not have substantial in-depth combustion.[27, 28] This correlation was compared with the flame height results of the source fires and the instrumented wall experiments.

FDS[22] and Smokeview[29] were used to examine the ability to numerically reproduce the source fires and the instrumented wall experiments.  Comparison between the experiments and the model will provide valuable validation data for the model.

@@

The natural gas burner was 30.5 cm on a side and the top surface of the burner was 7.5~cm above the floor. The shell of the burner was made from steel and filled with “pea sized” gravel with an average diameter of 6~mm.  A steady state heat release rate of approximately 80 kW was used for the experiments.

Gasoline was the fuel of choice for the liquid fueled pan fire.  Gasoline is not an ideal fuel to use in experiments because it is a mixture of components and additives that are specific to manufacturers. In addition the mixtures change from season to season, and vary based on requirements or restrictions of the locality.  However, fire incident data shows that gasoline was the first item ignited in the majority of intentional fires where flammable or combustible liquids were used 12.[30] In addition, data from forensic laboratories, collected by Babrauskas and others, indicates that gasoline is the most prevalent accelerant found as part of the analysis of fire debris samples 13, 14.[31, 32]
The gasoline pan fire used 500 mL of unleaded, 87 octane gasoline.  The pan was constructed from 6~mm thick steel.  The inner dimension of the pan was 30.5 cm on a side and 1.9~cm high.  The pan was elevated in order to bring the pan lip height up to 7.5~cm above the floor to be the same height as the natural gas burner.  The 500~mL of gasoline had a fuel depth of 5.5~mm and provided a burn time of approximately 300~s.

Polyurethane foam was chosen as the solid fuel.  According to the Polyurethane Foam Association, flexible polyurethane foam is the most widely used cushioning material in upholstered furnishing and mattresses. More than 1.2 billion pounds are produced in the U.S. every year 15,16.[33, 34] In studies by Ahrens and Evarts of the NFPA, upholstered furniture and mattresses and bedding were the first items ignited in 33~\% of the fatal fires in the U.S. based on averaged data from 2005 through 2009.[35, 36]

>>>>

Heskestad developed an empirical method for calculating the mean visible flame height for fuels which do not have substantial in-depth combustion~\cite{Beyler:1986,Heskestad:SFPE}. This correlation was compared with the flame height results of the source fires and the instrumented wall experiments.

FDS~\cite{McGrattan:2014} and Smokeview~\cite{Forney:2014} were used to examine the ability to numerically reproduce the source fires and the instrumented wall experiments.  Comparison between the experiments and the model will provide valuable validation data for the model.

@@

The natural gas burner was 30.5 cm on a side and the top surface of the burner was 7.5~cm above the floor. The shell of the burner was made from steel and filled with ``pea sized'' gravel with an average diameter of 6~mm.  A steady state heat release rate of approximately 80 kW was used for the experiments.

Gasoline was the fuel of choice for the liquid fueled pan fire.  Gasoline is not an ideal fuel to use in experiments because it is a mixture of components and additives that are specific to manufacturers. In addition the mixtures change from season to season, and vary based on requirements or restrictions of the locality.  However, fire incident data shows that gasoline was the first item ignited in the majority of intentional fires where flammable or combustible liquids were used~\cite{Rohr:2001}. In addition, data from forensic laboratories, collected by Babrauskas and others, indicates that gasoline is the most prevalent accelerant found as part of the analysis of fire debris samples~\cite{Babrauskas:2003,Chasteen:2010}.

The gasoline pan fire used 500~mL of unleaded, 87 octane gasoline.  The pan was constructed from 6~mm thick steel.  The inner dimension of the pan was 30.5~cm on a side and 1.9~cm high.  The pan was elevated in order to bring the pan lip height up to 7.5~cm above the floor to be the same height as the natural gas burner.  The 500~mL of gasoline had a fuel depth of 5.5~mm and provided a burn time of approximately 300~s.

Polyurethane foam was chosen as the solid fuel.  According to the Polyurethane Foam Association, flexible polyurethane foam is the most widely used cushioning material in upholstered furnishing and mattresses. More than 1.2 billion pounds are produced in the U.S. every year~\cite{Foam,Touch}. In studies by Ahrens and Evarts of the NFPA, upholstered furniture and mattresses and bedding were the first items ignited in 33~\% of the fatal fires in the U.S. based on averaged data from 2005 through 2009~\cite{Ahrens:2011,Evarts:2011}.
